Forecasting future oil production in Norway and the UK: a general improved methodology

### 2014-08-14 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

\usepackage{abstract}% Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

@@

@@

\maketitle
\begin{abstract}
\noindent We present a new Monte-Carlo methodology to forecast
the crude oil production of Norway and the U.K. based on a two-step process,
(i) the nonlinear extrapolation of the current/past
performances of individual oil fields and (ii) a stochastic model
of the frequency of future oil field discoveries. Compared
with the standard methodology that tends to underestimate remaining oil reserves,
our method gives a better description of future oil production, as validated by
our back-tests starting in 2008. Specifically, we predict remaining reserves
extractable until 2030 to be $5.7 \pm 0.3$ billion barrels for Norway and $3.0 \pm0.3$ billion barrels for the UK, which are respectively $45\%$ and $66\%$ above the predictions using an extrapolation of aggregate production.
\end{abstract}

\section{Introduction}

\lettrine[nindent=0em,lines=3] Forecasting future oil production
has been a topic of active interest since the beginning of the past
century because of oil central role in our economy. Its importance
ranges from energy production, manufacturing to the pharmaceutical
industry. As petroleum is a non-renewable and finite resource, it
is primordial to be able to forecast future oil production. The fear
of a global oil peak, beyond which production will inevitably decline,
has been growing due to stagnating supplies and high oil prices since
the crisis in 2008/2009 \citep{Murray2013}. As any industrialized
country, Europe is strongly dependent on oil supply to maintain its
economic power. In the nowadays difficult geopolitical environment,
it is important to know how much of the oil needed in Europe will
come from reliable sources. In the past, a big share has been coming
from Norway and the U.K., two of Europe's biggest exporters. However,
the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
production has been declining rapidly as well \citep{Hook2008}.\\
 The methodology behind forecasting future oil production has not
evolved much since M. King Hubbert, who in 1956 famously predicted
that the U.S. oil production would peak around 1965-1970 \citep{Hubbert1956}.
That prediction has proven itself to be correct. His main argument
was based on the finiteness of oil reserves and to what amounts
as the use of the logistic differential equation for the total
quantity $P(t)$ of oil extracted up to time $t$
\begin{equation}
\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)~.
\label{logistic}
\end{equation}
The logistic differential equation
is characterized by an initial exponential growth, which then decreases
to zero as the total oil extracted reaches saturation (no more oil
is to be found). The parameter $r$ is commonly referred to as the
growth rate, and $K$ as the carrying capacity (total quantity of
oil that can be ultimately extracted). If $P(t)$ is the amount of oil extracted up to
time $t$, then $f(t) := \frac{dP}{dt}$ is the oil production rate, the quantity
that M. King Hubbert predicted with surprising accuracy to peak. From
a methodological point of view, the Hubbert model has enjoyed a longstanding
popularity in modeling future oil production given its simplicity.
Various extensions have been studied by \citet{Brandt2007}
to account for multi-cycled or asymmetric production curves.

@@

Despite the popularity of the Hubbert model, the literature says little about its out-of-sample performance. There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time. As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test had to be started when the peak is already well visible in the data. For each time $T$, a month between the start of the back-test and the date of the latest available date, we fitted the Hubbert model to the data truncated at time $T$. Based on each fit, we computed the predicted remaining recoverable reserves in 2014.\\
This back-test allowed us to study how the forecast of Hubbert's model evolves over time. A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time. An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects. This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.\\
For the studied cases, namely Norway and the UK, a good starting date was chosen to be the 01/2000. Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which required the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks. The peak of each of the two logistic functions is matched up with one of the two peaks in the UK production data.\\
The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time. In 2000, for Norway, the Hubbert model would have predicted 2.08Gb remaining recoverable reserves in 2014, while up to now the predicted amount grew to 2.83Gb (a 36\% error) and the growth did not yet slow down. For the UK, the prediction would have been 0.21Gb but grew all the way to 1.43Gb (an error of 581\%), with a recent slow down in growth. These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast. For completeness, it has to be added that the relative error between the year 2000 and now in total production since the beginning is only 17\% for Norway and 6.5\% for the UK. Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.\\
The aim is to understand the causes why the Hubbert model underestimates the tail of the production curve and improve upon it. An argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves. To discuss this hypothesis the figure \ref{hubbert-backtest} shows as well the oil price. The forecast for Norway remains fairly stable in the beginning and then starts to grow with a small delay when the oil price increases. The growth is continuing in the past years, even as the oil price stabilized again. For the UK the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in the oil price. One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to the oil price as new capacities had first to be build. On the other hand, UK could have had available or upcoming spare capacities which allow to respond rapidly to the increase in the oil price. While it is certain that increases in the oil price will increase the recoverable reservers, it seems difficult to establish a mathematically defined link between the oil price and the increase in forecasted reserves. Consequently, an extension of the Hubbert model integrating the oil price has not been further explored.

@@

The existing forecasts of future oil production all use some form of the
Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
or some economical model applied to aggregate production \citep{Greiner2011}.
 However, as just shown in the back-test, the Hubbert model does not provide a
forecast stable in time, usually underestimating the tail in the production curve.
Considering that none of these Hubbert based methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data. The Hubbert model will certainly always remain the tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.

@@

A comparison with the results in table 1 of \cite{Smith1980} show that our results a coherent. Back in 1980, he used $J=7$ which fortunately maps easily to our splitting using $J=2$ as his first reservoir size is 0 to 50 Mb as ours, and the sum of the 6 others yields our second reservoir size 50 Mb and above. His estimated total reserves were 43.2 billion barrels with 90\% confidence bounds at 38.4 and 65.0, across 203 dwarfs and 117 giants in our definition. Currently, we know that total estimated reserves are reaching 59 Bb, which is still within the 90\% confidence interval of Smith. This strongly speaks in favour of his likelihood technique. Given that in the past 34 years there have been major technological improvements and a strong increase in the oil price, it is not surprising that the actual reserves are tending towards his upper bound.\\
A linear scaling by 59/43.2 (actual reserves over Smith expected reserves) of the number of fields Smith predicted, yields 277 dwarfs and 160 giants. This is close to the $296 \pm 21$ dwarfs and $156 \pm 2$ giants our 60\% likelihood estimation found for Norway and the UK combined. Such a strong coherence between the two fully independent implementations using different parameters (e.g. $J=7$ versus $J=2$) and different data sets (1980 versus 2014) is almost surprising and we see it as a good validation of the approach.

@@

The results in table \ref{forecast-14} show a striking difference
between the Hubbert model, the extrapolation of the fit and the Monte-Carlo model forecast. The preliminary study discussed in the introduction already showed that the Hubbert model forecast was not stable over time and therefore unlikely to be accurate. Indeed the results confirm that the Hubbert forecast is smaller then fitting the decaying process and only half of what the Monte-Carlo model predicts. This is not surprising, because the Hubbert curve is symmetric and therefore the fitting procedure will be strongly biased by the rapid increase in production before the peak, therefore predicting a rapid decrease. Only fitting the decay process using a stretched exponential resolves this problem, because the decay is entirely decoupled from the increase before the peak. Consequently, we expect the simple fit of the decay to perform better.\\
According to the fit (extrapolation of aggregate production), Norway's future oil production would still decay much faster than in the Monte-Carlo case. The remaining reserves estimated with the Monte-Carlo methodology are $45\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue to fit a double logistic (Hubbert model) or extrapolate the decay of aggregate production starting at the second peak. The differences between the three methodologies are very similar to the Norwegian case. The Hubbert model again predicts the smallest remaining reserves with about half the amount of the Monte-Carlo model. The fit of aggregate production underestimates the remaining oil reserves by about $66\%$ compared to the Monte-Carlo model.

@@

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the resulting production curves for the fit of aggregate production and the Monte-Carlo model.\\
Comparing the forecast of the models with the oil production of the subsequent 6 years, shows that the predictive power of the Monte-Carlo model is significantly better than the Hubbert model or a simple extrapolation of the aggregate past production. Table \ref{backtest-08} summarizes the difference between the approaches for the back-testing period. The Hubbert model has the poorest performance, underestimating production by 19\% (Norway) and 47\% (UK). The Monte-Carlo model is found essentially on target with less then 5\% error on the actual production, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K. These results confirm that modelling the production dynamics of the individual fields and the discovery of new fields yields a more accurate prediction then methodologies who just use aggregate production.\\
The error of the Hubbert model or the extrapolation of the past oil production method (``fit'') is not dramatic over the six years from 2008 to 2014, however the difference between them and the Monte-Carlo approach becomes huge for the time period from 2014 to 2030, as shown in table \ref{forecast-08}. The Hubbert model and the simple extrapolation decay too fast and entirely miss the fat tails in the decay process of individual fields and the new discoveries. It is interesting to note that the Hubbert model performs similarly then the fit of aggregate production for the UK, but performs much better for Norway. This is due to the fact that the Hubbert model calibrates the decay slope based on the increase before the peak, which in the case of Norway turns out to be beneficial.\\
The simple extrapolation changed massively between the back-test in table \ref{forecast-08} and the current fit in table \ref{forecast-14} (520\% for Norway and 236\% for the U.K.). In other words, the simple extrapolation is very unstable in its forecast, while the Monte-Carlo forecasts remains very consistent (less than 10\% change).\\
As can be seen in figure \ref{model-back-testing}, the actual production
of Norway during the back-testing period remained entirely within
the quite narrow $1\sigma$ interval of the Monte-Carlo methodology,
while totally breaking out of the $1\sigma$ interval of the simple
extrapolation. For the U.K. the Monte-Carlo methodology only performs
slightly better when considering the confidence interval, and the
confidence interval is much larger due to the uncertainty on future
discoveries and their production profile.

@@

The detailed discussion of a full back-test of the Hubbert model for Norway and the UK showed that its prediction were unstable in time and therefore unlikely to be reliable. We argue that the Hubbert model is likely to be unable to correctly capture the fat tail in the decay process as it doesn't model the fat tails in the decay of individual fields nor the discovery of new fields. An easily quantifiable link with the oil price, which is the driving factor of oil production, could not be found and we therefore rejected the exploration of extensions of the Hubbert model to capture properly the tail.\\
We preferred a two step Monte-Carlo based methodology to forecast future oil production which: (i) extends the oil production of current fields into the future and (ii) models the discovery rate of new fields. This methodology was then applied to forecast the future oil production of Norway and the U.K. These forecasts are significantly different from the ones obtained with an extrapolation of aggregate production. Indeed, our model forecasts 45\% to 66\% more remaining oil reserves than the standard extrapolation. The back-test performed on the time period between 2008 and 2014 confirmed that the Monte-Carlo based model better captured the production dynamics.

@@

%FIGURE 1
\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{oseberg}
\caption{Example of the time evolution of the oil production per day of
a regular field, parametrised beyond the peak in the decay
regime by the stretched exponential 
function (\ref{stretched_exponential}) with $\beta=0.66\pm0.01$ and $\tau=-55\pm1$ months,
shown with the black line.  The one standard deviation given by expression
(\ref{eq:extrapolation}) is represented by the grey band.}
\label{regular} 
\end{figure}

@@

%FIGURE 4
\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{oesberg-error}
\caption{Oesberg field - Relative error defined by expression (\ref{ryjryukoik})
of the predicted total production from time $t$ indicated in the abscissa until 2014.
One can observe that the predicted future total production is over-estimated by as much
as 70\% in 1999, then under-estimated by the same amount in 2002, while the forecast
errors remain smaller than 20\% since 2004.}
\label{oesberg-error}
\end{figure}

%FIGURE 5
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
\caption{Logistic fit of the function solution of expression (\ref{logistic})
to the number of discoveries for Norway. The discovery
rate of new oil fields is dependent on their size as explained in the main text.}
\label{rate_of_discovery} 
\end{figure}

@@

%FIGURE 6
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{ccdf-field-size}
\caption{Complementary 
cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
The two dashes lines visualise the power law behaviour (\ref{ccdfsizefield}) of the tail
of the distributions. }
\label{ccdffieldsizes}
\end{figure}

@@

%FIGURE 7
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{no-returns-size}
\caption{Total estimated revenue divided by investment as a function of the estimated 
ultimate recovery for a number of small oil fields. }
\label{returnNorwayfield}
\end{figure}

%FIGURE 8
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{NO-future}
\includegraphics[width=1\columnwidth]{UK-future} 
\caption{Monte-Carlo (green upper
continuous line with standard deviation bank starting in 2014 onward)
and fit forecast based on past production data (lower line and grey one standard deviation band) for Norway
(top) and the U.K. (bottom). In both cases, the Monte-Carlo model forecasts
a significantly slower decay than the fit by taking into account that
new fields will come in production.}
\label{no_uk_2013} 
\end{figure}

%FIGURE 9
\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{NO-back-test}
\includegraphics[width=1\columnwidth]{UK-back-test} \caption{Monte-Carlo (green upper
continuous line with standard deviation bank starting in 2014 onward) and Hubbert forecast based 
on past production data up to 2008 (lower line and grey one standard deviation band)
for Norway (top) and the U.K. (bottom). The results can be
compared with the subsequent oil production (blue area). In both cases,
the Monte-Carlo methodology is more precise.}
\label{model-back-testing} 
\end{figure}

@@

\begin{table}[h]
\caption{Likelihood estimation for the number of dwarf (1) and giant (2) fields.
For Norway, our logistic fit suggests that up to two new giant fields could be
ultimately discovered. For the U.K., the prediction is more bleak, suggesting
that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
\centering %
\begin{tabular}{lllll}
 & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{likelihood-results}
\end{table}

@@

\begin{table}[H]
\caption{Extrapolation of past oil production (``fit'') and prediction
using the Monte-Carlo model are made on the data set truncated
in 2008. Their forecast for the period 2008-2014 is compared to
the actual realised production. All units in billion barrels (Gb).}
\centering %
\begin{tabular}{lllll}
 & Actual (Gb) & Hubbert (Gb) & Fit (Gb) & Model (Gb)\tabularnewline
\midrule Norway & $4.05$  & $3.29$  & $3.28$  & $3.95$ \tabularnewline
U.K. & $2.40$  & $1.63$  & $2.00$  & $2.28$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}

>>>>

\usepackage{abstract}% Allows abstract customization
\usepackage{setspace}
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

@@

\doublespacing

@@

\maketitle
\begin{abstract}
\doublespacing
\noindent We present a new Monte-Carlo methodology to forecast
the crude oil production of Norway and the U.K. based on a two-step process,
(i) the nonlinear extrapolation of the current/past
performances of individual oil fields and (ii) a stochastic model
of the frequency of future oil field discoveries. Compared
with the standard methodology that tends to underestimate remaining oil reserves,
our method gives a better description of future oil production, as validated by
our back-tests starting in 2008. Specifically, we predict remaining reserves
extractable until 2030 to be $5.7 \pm 0.3$ billion barrels for Norway and $3.0 \pm0.3$ billion barrels for the UK, which are respectively $45\%$ and $66\%$ above the predictions using an extrapolation of aggregate production.
\end{abstract}

\lettrine[nindent=0em,lines=3] Forecasting future oil production
has been a topic of active interest since the beginning of the past
century because of oil central role in our economy. Its importance
ranges from energy production, manufacturing to the pharmaceutical
industry. As petroleum is a non-renewable and finite resource, it
is primordial to be able to forecast future oil production. The fear
of a global oil peak, beyond which production will inevitably decline,
has been growing due to stagnating supplies and high oil prices since
the crisis in 2008/2009 \citep{Murray2013}. As any industrialized
country, Europe is strongly dependent on oil supply to maintain its
economic power. In the nowadays difficult geopolitical environment,
it is important to know how much of the oil needed in Europe will
come from reliable sources. In the past, a big share has been coming
from Norway and the U.K., two of Europe's biggest exporters. However,
the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
production has been declining rapidly as well \citep{Hook2008}.
This motivates the novel Monte-Carlo methodology to forecast
the crude oil production of Norway and the U.K. that we present here.

\section{State of the art}

\subsection{Hubbert's Model for peak oil forecasting}

The methodology behind forecasting future oil production has not
evolved much since M. King Hubbert, who in 1956 famously predicted
that the U.S. oil production would peak around 1965-1970 \citep{Hubbert1956}.
That prediction has proven itself to be correct. His main argument
was based on the finiteness of oil reserves and to what amounts
as the use of the logistic differential equation for the total
quantity $P(t)$ of oil extracted up to time $t$
\begin{equation}
\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)~.
\label{logistic}
\end{equation}
The logistic differential equation
is characterized by an initial exponential growth, which then decreases
to zero as the total oil extracted reaches saturation (no more oil
is to be found). The parameter $r$ is commonly referred to as the
growth rate, and $K$ as the carrying capacity (total quantity of
oil that can be ultimately extracted). If $P(t)$ is the amount of oil extracted up to
time $t$, then $f(t) := \frac{dP}{dt}$ is the oil production rate, the quantity
that M. King Hubbert predicted with surprising accuracy to peak. From
a methodological point of view, the Hubbert model has enjoyed a longstanding
popularity in modeling future oil production given its simplicity.
Various extensions have been studied by \citet{Brandt2007}
to account for multi-cycled or asymmetric production curves.

@@

Despite the popularity of the Hubbert model, the literature says little about its out-of-sample performance,
apart from the one case that made the prediction famous, namely the prediction by Hubbert of the
 conventional oil peak that occurred in the US in 1970. There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time. As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test had to be started when the peak is already well visible in the data. For each time $T$, a month between the start of the back-test and the date of the latest available date, we fitted the Hubbert model to the data truncated at time $T$. Based on each fit, we computed the predicted remaining recoverable reserves in 2014.\\
This back-test allowed us to study how the forecast of Hubbert's model evolves over time. A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time. An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects (technically, it is called ``misspecified''). This is of course assuming that the dynamics of the system is deterministic to a higher degree than the error the model is making in its forecast.\\
For the studied cases, namely Norway and the UK, a good starting date was chosen to be 01/2000. Fitting Norway's production could be made using the simple logistic equation \ref{logistic}. In contrast, for the UK, the production exhibits two clear peaks, which required the use of a double-logistic function that is the sum of two independent logistic functions with time shifted peaks. The peak of each of the two logistic functions is matched up with one of the two peaks in the UK production data.\\
The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time. In 2000, for Norway, the Hubbert model would have predicted 2.08Gb of remaining recoverable reserves by 2014 while, up to the time of writing (July 2014), the predicted amount grew to 2.83Gb (a 36\% error) and the growth is not showing sign of slowing down, yet. For the UK, the prediction would have been 0.21Gb but grew all the way to 1.43Gb (an error of 581\%), with a recent slow down in growth. These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast of recoverable reserves. For completeness, let us stress that the relative error between the year 2000 and July 2014 in total production since the beginning is only 17\% for Norway and 6.5\% for the UK. Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.\\
Our aim is first to understand why the Hubbert model underestimates the tail of the production curve and then to improve upon it. An argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves. To discuss this hypothesis, figure \ref{hubbert-backtest} shows the oil price as well. The forecast for Norway remains fairly stable from 2000 to 2007 during a period when the oil
price went from slightly above \$20 per barrel to more than \$80 per barrel. The shift to a larger predicted
remaining production does not correlate with the peak, bottom and then rebound of the oil prices
since 2007. For the UK, the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in oil prices. One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to oil prices as new capacities had first to be build. On the other hand, UK could have had available or upcoming spare capacities that allowed to respond rapidly to the increase in the oil price. While it is certain that increases in oil prices will augment the recoverable reserves, it seems difficult to establish a well-defined mathematical link between oil price and the increase in forecasted reserves. Consequently, an extension of the Hubbert model integrating the oil price has not been further explored.

@@

The existing forecasts of future oil production all use some version of the
Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
or some economical model applied to aggregate production \citep{Greiner2011}.
 However, as just shown in the back-test, the Hubbert model does not provide a
forecast that is stable in time, usually underestimating the significant tail in the production curve.
Considering that none of these Hubbert-based methodologies go into the details of studying the underlying dynamics, it is not surprising that important features of the dynamics bearing
on the total recoverable reserves have been missed.
In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data. For this reason, the Hubbert model will likely remain a tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.

@@

A comparison with the results in table 1 of \cite{Smith1980} show that our results are coherent. Back in 1980, \cite{Smith1980} used $J=7$, which fortunately maps easily to our splitting using $J=2$ as his first reservoir size is 0 to 50 Mb as ours, and the sum of the 6 others yields our second reservoir size 50 Mb and above. His estimated total reserves were 43.2 billion barrels with 90\% confidence bounds at 38.4 and 65.0, across 203 dwarfs and 117 giants in our definition. Currently, we know that total estimated reserves are reaching 59 Bb, which is still within the 90\% confidence interval of Smith. This strongly speaks in favour of his likelihood technique. Given that, in the past 34 years, there have been major technological improvements and a strong increase in the oil price, it is not surprising that the actual reserves are tending towards his upper bound.\\
A linear scaling by 59/43.2 (actual reserves over Smith expected reserves) of the number of fields Smith predicted, yields 277 dwarfs and 160 giants. This is close to the $296 \pm 21$ dwarfs and $156 \pm 2$ giants of our 60\% likelihood estimation found for Norway and the UK combined. Such a strong coherence between the two fully independent implementations using different parameters (e.g. $J=7$ versus $J=2$) and different data sets (1980 versus 2014) provides a credible validation of the approach.

@@

The results in table \ref{forecast-14} show a striking difference
between the Hubbert model, the extrapolation of the fit and the Monte-Carlo model forecast. The preliminary study discussed in the introduction already showed that the Hubbert model forecast was not stable over time and therefore unlikely to be accurate. Indeed the results confirm that the Hubbert forecast is smaller then fitting the decaying process and only half of what the Monte-Carlo model predicts. This is not surprising, because the Hubbert curve is symmetric and therefore the fitting procedure will be strongly biased by the rapid increase in production before the peak, therefore predicting a rapid decrease. Only fitting the decay process using a stretched exponential resolves this problem, because the decay is entirely decoupled from the increase before the peak. Consequently, we expect the simple fit of the decay to perform better.\\
According to the fit (extrapolation of aggregate production), Norway's future oil production would still decay much faster than in the Monte-Carlo case. The remaining reserves estimated with the Monte-Carlo methodology are $45\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue as this inverted \textquotedbl{}w shape\textquotedbl{}
can be fitted well by a double logistic (Hubbert model). Alternatively, we can extrapolate the decay of the aggregate production starting at the second peak. The differences between the three methodologies are very similar to the Norwegian case. The Hubbert model again predicts the smallest remaining reserves with about half the amount of the Monte-Carlo model. The fit of the aggregate production underestimates the remaining oil reserves by about $66\%$ compared to the Monte-Carlo model.

@@

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the resulting production curves for the fit of the aggregate production and of the Monte-Carlo model.\\
Comparing the forecast of the models with the oil production of the subsequent 6 years shows that the predictive power of the Monte-Carlo model is significantly better than the Hubbert model or than a simple extrapolation of the aggregate past production. Table \ref{backtest-08} summarizes the difference between the approaches for the back-testing period. The Hubbert model has the poorest performance, underestimating production by 19\% (Norway) and 47\% (UK). The Monte-Carlo model is found essentially on target with less then 5\% error compared with the actual production, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K. These results confirm that modelling the production dynamics of the individual fields and the discovery of new fields yields a more accurate prediction then methodologies that just use aggregate production.\\
The errors of the Hubbert model and of the extrapolation of the past oil production method (``fit'') are not dramatic over the six years from 2008 to 2014. However, the difference between them and the Monte-Carlo approach becomes huge for the time period from 2014 to 2030, as shown in table \ref{forecast-08}. The Hubbert model and the simple extrapolation decay too fast and entirely miss the fat tails in the decay process of individual fields and the new discoveries. It is interesting to note that the Hubbert model performs similarly then the fit of aggregate production for the UK, but performs much better for Norway. This is due to the fact that the Hubbert model calibrates the decay slope based on the increase before the peak, which in the case of Norway turns out to be beneficial.\\
The simple extrapolation changed massively between the back-test in table \ref{forecast-08} and the current fit in table \ref{forecast-14} (520\% for Norway and 236\% for the U.K.). In other words, the simple extrapolation is very unstable in its forecast, while the Monte-Carlo forecasts remains very consistent (less than 10\% change).\\
As can be seen in figure \ref{model-back-testing}, the actual production
of Norway during the back-testing period remained entirely within
the quite narrow $1\sigma$ interval of the Monte-Carlo methodology,
while totally breaking out of the $1\sigma$ interval of the simple
extrapolation. For the U.K. the Monte-Carlo methodology only performs
slightly better when considering the confidence interval, and the
confidence interval is much larger due to the uncertainty on future
discoveries and their production profile.

@@

The detailed discussion of a full back-test of the Hubbert model for Norway and the UK showed that its prediction were unstable in time and therefore unlikely to be reliable. We argue that the Hubbert model is likely to be unable to correctly capture the fat tail in the decay process as it does not model the fat tails in the decay of individual fields nor the discovery of new fields. An easily quantifiable link with the oil price, which is the driving factor of oil production, could not be found and we therefore did not pursue the exploration of extensions of the Hubbert model to capture properly the tails of the decay production process.\\
We preferred a two step Monte-Carlo based methodology to forecast future oil production which: (i) extends the oil production of current fields into the future and (ii) models the discovery rate of new fields. This methodology was then applied to forecast the future oil production of Norway and the U.K. These forecasts are significantly different from the ones obtained with an extrapolation of aggregate production. Indeed, our model forecasts 45\% to 66\% more remaining oil reserves than the standard extrapolation. The back-test performed on the time period between 2008 and 2014 confirmed that the Monte-Carlo based model better captured the production dynamics.

@@

%FIGURE 1
\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{oseberg}
\caption{\doublespacing
Example of the time evolution of the oil production per day of
a regular field, parametrised beyond the peak in the decay
regime by the stretched exponential 
function (\ref{stretched_exponential}) with $\beta=0.66\pm0.01$ and $\tau=-55\pm1$ months,
shown with the black line.  The one standard deviation given by expression
(\ref{eq:extrapolation}) is represented by the grey band.}
\label{regular} 
\end{figure}

@@

%FIGURE 4
\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{oesberg-error}
\caption{\doublespacing
Oesberg field - Relative error defined by expression (\ref{ryjryukoik})
of the predicted total production from time $t$ indicated in the abscissa until 2014.
One can observe that the predicted future total production is over-estimated by as much
as 70\% in 1999, then under-estimated by the same amount in 2002, while the forecast
errors remain smaller than 20\% since 2004.}
\label{oesberg-error}
\end{figure}

%FIGURE 5
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
\caption{\doublespacing
Logistic fit of the function solution of expression (\ref{logistic})
to the number of discoveries for Norway. The discovery
rate of new oil fields is dependent on their size as explained in the main text.}
\label{rate_of_discovery} 
\end{figure}

@@

%FIGURE 6
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{ccdf-field-size}
\caption{\doublespacing
Complementary cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
The two dashes lines visualise the power law behaviour (\ref{ccdfsizefield}) of the tail
of the distributions. }
\label{ccdffieldsizes}
\end{figure}

@@

%FIGURE 7
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{no-returns-size}
\caption{\doublespacing
Total estimated revenue divided by investment as a function of the estimated 
ultimate recovery for a number of small oil fields. }
\label{returnNorwayfield}
\end{figure}

%FIGURE 8
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{NO-future}
\includegraphics[width=1\columnwidth]{UK-future} 
\caption{\doublespacing
Monte-Carlo (green upper continuous line with standard deviation bank starting in 2014 onward) and fit forecast based on past production data (lower line and grey one standard deviation band) for Norway (top) and the U.K. (bottom). In both cases, the Monte-Carlo model forecasts a significantly slower decay than the fit by taking into account that new fields will come in production.}
\label{no_uk_2013} 
\end{figure}

%FIGURE 9
\noindent 
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{NO-back-test}
\includegraphics[width=1\columnwidth]{UK-back-test}
\caption{\doublespacing
Monte-Carlo (green upper continuous line with standard deviation bank starting in 2014 onward) and Hubbert forecast based on past production data up to 2008 (lower line and grey one standard deviation band) for Norway (top) and the U.K. (bottom). The results can be compared with the subsequent oil production (blue area). In both cases, the Monte-Carlo methodology is more precise.}
\label{model-back-testing} 
\end{figure}

@@

\begin{table}[h]
\caption{\doublespacing
Likelihood estimation for the number of dwarf (1) and giant (2) fields. For Norway, our logistic fit suggests that up to two new giant fields could be ultimately discovered. For the U.K., the prediction is more bleak, suggesting that the most probable scenario is that one giant field will be found (which is most likely already discovered but classified as a new field).}
\centering %
\begin{tabular}{lllll}
 & $m_{N1}$ & $n_{1}\pm\sigma_{1}$ & $m_{N2}$ & $n_{2}\pm\sigma_{2}$\tabularnewline
\midrule Norway  & $24$ & $88.4\pm10.0$ & $52$ & $56.4\pm1.6$\tabularnewline
U.K. & $162$ & $208\pm11$ & $99$ & $100\pm0.4$\tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{likelihood-results}
\end{table}

@@

\begin{table}[H]
\caption{\doublespacing
Extrapolation of past oil production (``fit'') and prediction
using the Monte-Carlo model are made on the data set truncated
in 2008. Their forecast for the period 2008-2014 is compared to
the actual realised production. All units in billion barrels (Gb).}
\centering %
\begin{tabular}{lllll}
 & Actual (Gb) & Hubbert (Gb) & Fit (Gb) & Model (Gb)\tabularnewline
\midrule Norway & $4.05$  & $3.29$  & $3.28$  & $3.95$ \tabularnewline
U.K. & $2.40$  & $1.63$  & $2.00$  & $2.28$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}
